---
title: "Project Part 1" 
author: ""
fontsize: 12pt
geometry: margin=1in
urlcolor: black
output: pdf_document
header-includes:
- \usepackage{setspace}
- \onehalfspacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, highlight=FALSE)
```

```{r, echo=FALSE}
library(carData)
library(ggplot2)
library(tidyverse)
library(gridExtra)
#Data for project report
head(Arrests)
```
## Data Set Description  
### Background information
This crime data is a subset of a larger dataset that was collected by the 
Toronto Star newspaper. It shows data from 1997-2002 on arrests for Marijuana 
possession in Toronto, Canada. The Toronto star newspaper collected this data to
show the differing levels of the city's police department's treatment towards
arrestees. The Star used the data to conduct an in-depth investigation of the
differences in treatment of blacks and whites by the Toronto police. 
The data was used by the newspaper to argue that blacks arrested by Toronto 
police are treated more harshly than whites. (6)

### Rows and Variables Explanation
The columns that will be described are only the ones that are relevant
to the graphical representation and to the conclusions that can be drawn from
it. Checks are the number of times in which the arrestee's name appeared in other
police department databases (through previous arrests, previous convictions, 
parole status, etc.) and shows the criminal history of an arrestee which could 
help with identifying differing levels of treatment by the Toronto police 
for blacks and whites. The year column specifies the year when the arrestee was 
arrested for marijuana possession and can help solidify the trend of levels of 
treatment for blacks and whites by the police department. The sex and colour 
columns specifies the sex and the color of the arrestee which is absolutely 
needed for showcasing differing levels of treatment for the two groups. 
The citizens column records the citizen status of the arrestee in Canada. 
The employed column records the employment status of the arrestee and attempts 
to go into detail about the economic circumstances of the arrestee.

### Data Collection Method
The Toronto Star newspaper collected its data from 
Toronto's police database which stores records of all of the incidents of its 
arrests, ticketing, etc for all of its arrestees along with some of their
background information. This data is a sample of the arrestees for
marijuana possession from 1997-2002. The data was sampled by Michael
Friendly from York University who randomly selected 1000 arrestees 
from the data that was gathered from Toronto Star newspaper 
and therefore a simple random sample was conducted to gather this data.
(6)

### Potential Issues
A potential issue with this data is that it doesn't exactly go into greater 
detail for the economic conditions for the arrestees. Employed individuals
who are in lower economic classes, because they're not paid much money 
in their job, may not be able to make enough money to support themselves and 
their family and therefore may have to turn to crime in order to
be able to make money for themselves. One of these ways could even be through 
dealing drugs which could result in frequent arrests for marijuna possession. 
Additionally, another issue with this data is that non-citizens of Canada 
who are arrested will probably not show accurate records of their previous 
arrests. Toronto's police database can only collect data on individuals 
who were arrested by Toronto's police and they don't have access to the police 
databases of other countries. Therefore, the number of checks for non-citizens 
of could be lower than they really should be, as some non-citizens would've been
arrested for crimes in another country, but may have only been
arrested once in Canada and only that database would show that individual's 
name.

## Conclusions
Overall, the data shows that black males have the highest average number of
checks from 1997-2002. This can be shown by the fact that all of the bars 
from 1997-2002 for the black males bar graph are higher than than 
the corresponding bars for the other bar graphs. Some of the factors that may
lead to this data could be the economic conditions for black males, or 
it may be the case that the Toronto police department is more harsh towards
black males and therefore this could lead to more arrests.
Furthermore, the data also shows that black females
had a higher average number of checks than white females from 1997-2002. This
can be shown by the fact that all of the bars from 1997-2002 for the black
females are higher than the corresponding bars for the white females. Some of
the factors that could lead to this data are the economic conditions for black
females, or that the police are more harsh towards blacks females than white
females which could lead to more arrests for them. (6)

## Data Set Graphical and Numerical Summaries
```{r}
arrestData <- data.frame(Arrests)
whiteMaleData <- data.frame(filter(arrestData, colour=="White" & sex =="Male"))
whiteFemaleData <- data.frame(filter(arrestData, colour == "White" & sex 
== "Female"))
blackMaleData <- data.frame(filter(arrestData, colour == "Black" & sex 
== "Male"))
blackFemaleData <- data.frame(filter(arrestData, colour == "Black" & sex 
== "Female"))

whiteMaleBar <- ggplot(whiteMaleData, aes(x=factor(year), y=checks))+
stat_summary(fun="mean", geom="bar", color="blue", fill = "white" )+
theme_bw()+labs(title="White Males", y = "Average Number of Checks", x = "year")+
theme(axis.title.y = element_text(size=8),  plot.title=
element_text(hjust=0.5, size=11))+expand_limits(y=c(0,2.5)) #(1),(4),(5)

whiteFemaleBar <- ggplot(whiteFemaleData, aes(x=factor(year), y=checks))+
stat_summary(fun="mean", geom="bar", color="pink", fill="white")+theme_bw()+
labs(title="White Females", y = "Average Number of Checks", x= "year")+
theme(axis.title.y = element_text(size=8), plot.title=
element_text(hjust=0.5, size=11))+expand_limits(y=c(0,2.5)) #(1),(4),(5)

blackMaleBar <- ggplot(blackMaleData, aes(x=factor(year), y=checks))+
stat_summary(fun="mean", geom="bar", color="blue", fill="black")+theme_bw()+
labs(title="Black Males", y = "Average Number of Checks", x="year")+
theme(axis.title.y = element_text(size=8), plot.title=
element_text(hjust=0.5, size=11))+expand_limits(y=c(0,2.5)) #(1),(4),(5)

blackFemaleBar <- ggplot(blackFemaleData, aes(x=factor(year), y=checks))+
stat_summary(fun="mean", geom="bar", color="pink", fill="black")+theme_bw()+
labs(title="Black Females", y = "Average Number of Checks", x= "year")+
theme(axis.title.y = element_text(size=8), plot.title=
element_text(hjust=0.5, size=11))+expand_limits(y=c(0,2.5)) #(1),(4),(5) 

#(2), (3)
grid.arrange(blackMaleBar, blackFemaleBar,whiteMaleBar, whiteFemaleBar, 
top = "Marijuna Possession Arrests in Toronto: Average Number of Checks") 
```
## References

1)<https://groups.google.com/forum/#!topic/ggplot2/Sz1lMjOI5JA>

2)<https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html>

3)<https://stackoverflow.com/questions/14726078/changing-title-in-multiplot-ggplot2-using-grid-arrange>

4)<https://rstudio-pubs-static.s3.amazonaws.com/366989_b80cfe1ffa5149f589fb44ea847f5967.html>

5)<http://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations>

6)<https://www.thestar.com/news/gta/knowntopolice/2002/10/19/singled-out.html>
